High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-04-05 notebooks/synthetic.qmd
more work on granular stuff

# Hyperparameters:
cvgs = [0.5, 0.75, 0.95]
temps = [0.01, 0.1, 1.0]
Λ = [0.0, 0.1, 1.0]
l2_λ = 0.1

@@

```{julia}
generators = Dict(
    :wachter => GenericGenerator(opt=opt, λ=l2_λ),
    :revise => REVISEGenerator(opt=opt, λ=l2_λ),
    :greedy => GreedyGenerator(opt=opt),
)

# Untrained Models:
models = Dict("cov$(cov)" => CCE.ConformalModel(conformal_model(mlp; method=:simple_inductive, coverage=cov)) for cov in cvgs)
```

```{julia}
bmks = []
measures = [
    CounterfactualExplanations.distance,
    CCE.distance_from_energy,
    CCE.distance_from_targets,
    CounterfactualExplanations.validity,
]
for (dataname, dataset) in datasets
    for λ in Λ, temp in temps
        _generators = deepcopy(generators)
        _generators[:cce] = CCEGenerator(temp=temp, λ=[l2_λ,λ], opt=opt)
        _generators[:energy] = CCE.EnergyDrivenGenerator(λ=[l2_λ,λ], opt=opt)
        _generators[:target] = CCE.TargetDrivenGenerator(λ=[l2_λ,λ], opt=opt)
        bmk = benchmark(
            dataset; 
            models=deepcopy(models), 
            generators=_generators, 
            measure=measures,
            suppress_training=false, dataname=dataname,
            n_individuals=5,
            initialization=:identity,
        )
        bmk.evaluation.λ .= λ
        bmk.evaluation.temperature .= temp
        push!(bmks, bmk)
    end
end
bmk = reduce(vcat, bmks)
```

@@

df = bmk()
for dataname ∈ sort(unique(df.dataname))
    Markdown.parse("""### $dataset""")
    df_ = df[df.dataname .== dataname, :]
    for λ in Λ, temp in temps
        Markdown.parse("""#### λ: $λ""")
        df_plot = df_[df_.λ .== λ, :]
        df_plot = df_plot[df_plot.temperature .== temp, :]
        plt = AlgebraOfGraphics.data(df_plot) * visual(BoxPlot) * 
            mapping(:generator, :value, row=:variable, col=:model, color=:generator)
        plt = draw(
            plt, axis=(xlabel="", xticksvisible=false, xticklabelsvisible=false, width=200, height=180), 
            facet=(; linkyaxes=:minimal)
        )   
        display(plt)
    end
end
```

>>>>

# Hyperparameters:
cvgs = [0.5, 0.75, 0.95]
temps = [0.01, 0.1, 1.0]
Λ = [0.0, 0.1, 1.0, 10.0]
l2_λ = 0.1

@@

```{julia}
# Benchmark generators:
generators = Dict(
    :wachter => GenericGenerator(opt=opt, λ=l2_λ),
    :revise => REVISEGenerator(opt=opt, λ=l2_λ),
    :greedy => GreedyGenerator(),
)

# Untrained Models:
models = Dict(Symbol("cov$(Int(100*cov))") => CCE.ConformalModel(conformal_model(mlp; method=:simple_inductive, coverage=cov)) for cov in cvgs)

# Measures:
measures = [
    CounterfactualExplanations.distance,
    CCE.distance_from_energy,
    CCE.distance_from_targets,
    CounterfactualExplanations.validity,
]
```

### Single CE

_temp = 0.01
results = DataFrame()
for (dataname, data) in datasets

    for (modelname, M) in deepcopy(models)

        # Model training:
        M = train(M, data)
        # Set up CCE:
        factual_label = predict_label(M, data, x)[1]
        target_label = data.y_levels[data.y_levels .!= factual_label][1]

        for λ in Λ

            # Generators:
            _generators = deepcopy(generators)
            _generators[:cce] = CCEGenerator(temp=_temp, λ=[l2_λ,λ], opt=opt)
            _generators[:energy] = CCE.EnergyDrivenGenerator(λ=[l2_λ,λ], opt=opt)
            _generators[:target] = CCE.TargetDrivenGenerator(λ=[l2_λ,λ], opt=opt)

            for (gen_name, gen) in _generators

                # CCE for given models, λ and generator:
                @assert predict_label(M, data, x) != target_label
                ce = try
                    generate_counterfactual(
                        x, target_label, data, M, gen;
                        initialization=:identity,
                        converge_when=:generator_conditions,
                        gradient_tol=gradient_tol,
                        max_iter=max_iter,
                    )
                catch
                    missing
                end

                if !ismissing(ce)
                    eval = DataFrame(evaluate(ce, measure=measures, output_format=:Dict))
                else
                    eval = DataFrame(Dict(Symbol(fun) => missing for fun in measures))
                end

                _results = DataFrame(
                    dataset = dataname,
                    model = modelname,
                    λ = λ,
                    generator = gen_name,
                    ce = ce,
                    factual = factual_label,
                    target = target_label,
                )

                _results = crossjoin(_results, eval; makeunique=true)
                append!(results, _results)

            end
        end
    end
end
```

function plot_benchmark(results; dataset=:multi_class, modelname=:cov95, img_height=300, zoom=-0.2)
    df_plot = results[results.dataset .== dataset,:] |>
        res -> res[res.model .== modelname,:]
    plts = map(eachrow(df_plot)) do row
        Plots.plot(
            row.ce,
            title="λ: $(row.λ), gen: $(row.generator)",
            cbar=false,
            zoom=zoom,
            legend=false,
        )
    end
    ncol = length(unique(df_plot.generator))
    nrow = length(unique(df_plot.λ))
    _layout = (nrow, ncol)
    Plots.plot(
        plts..., 
        size=img_height.*reverse(_layout), layout=_layout,
        plot_title="dataset: $dataset, model: $modelname",
    )
end
```

df = @pivot_longer(results, distance:distance_from_targets)
for dataname ∈ sort(unique(df.dataset))
    Markdown.parse("""### $dataname""")
    df_ = df[df.dataset .== dataname, :]
    for model in unique(df_.model)
        Markdown.parse("""#### model: $model""")
        df_plot = df_[df_.model .== model, :]
        df_plot = @mutate(df_plot, lambda = string("λ: ", round(λ, digits=2)))
        plt = AlgebraOfGraphics.data(df_plot) * visual(BarPlot) * 
            mapping(:generator, :value, row=:variable, col=:lambda, color=:generator)
        plt = draw(
            plt, axis=(xlabel="", xticksvisible=false, xticklabelsvisible=false, width=200, height=180), 
            facet=(; linkyaxes=:minimal)
        )   
        # plt.figure[0, :] = Label(
        #     plt.figure, "data: $dataname, model: $model", 
        #     fontsize=20, tellwidth=false
        # )
        display(plt)
    end
end
```

### Full Benchmark

```{julia}
bmks = []
for (dataname, dataset) in datasets
    for λ in Λ, temp in temps
        _generators = deepcopy(generators)
        _generators[:cce] = CCEGenerator(temp=temp, λ=[l2_λ,λ], opt=opt)
        _generators[:energy] = CCE.EnergyDrivenGenerator(λ=[l2_λ,λ], opt=opt)
        _generators[:target] = CCE.TargetDrivenGenerator(λ=[l2_λ,λ], opt=opt)
        bmk = benchmark(
            dataset; 
            models=deepcopy(models), 
            generators=_generators, 
            measure=measures,
            suppress_training=false, dataname=dataname,
            n_individuals=5,
            initialization=:identity,
        )
        bmk.evaluation.λ .= λ
        bmk.evaluation.temperature .= temp
        push!(bmks, bmk)
    end
end
bmk = reduce(vcat, bmks)
```

@@

df = bmk()
for dataname ∈ sort(unique(df.dataname))
    Markdown.parse("""### $dataname""")
    df_ = df[df.dataname .== dataname, :]
    for λ in Λ, temp in temps
        Markdown.parse("""#### λ: $λ""")
        df_plot = df_[df_.λ .== λ, :]
        df_plot = df_plot[df_plot.temperature .== temp, :]
        plt = AlgebraOfGraphics.data(df_plot) * visual(BoxPlot) * 
            mapping(:generator, :value, row=:variable, col=:model, color=:generator)
        plt = draw(
            plt, axis=(xlabel="", xticksvisible=false, xticklabelsvisible=false, width=200, height=180), 
            facet=(; linkyaxes=:minimal)
        )   
        display(plt)
    end
end
```
